---
title: "<span class='ex-title'>Spatial Analysis with R:</span><br/><span class='ex-subtitle'>Exercise 3: Import and Plot Spatial Data</span>"
date: "`r load('../common/course_info.RData'); lstCI$course`"
output: html_document
---

```{r setup, child = '../common/exercises_setup.Rmd', include=FALSE}
```

```{r preload_libraries, include=FALSE}
library(rgdal)
```

```{r instr, child = '../common/exercises_instructions.Rmd', include=FALSE}
```

<hr class="hrdivider">
<div class="copy-button"><button onclick="CopyToClipboard('qtn_div');return false;">Copy questions</button></div>
<div id="qtn_div" class="ex-qtn">

`r hashtitle("IMPORT A SHAPEFILE")`

<a id="qtn_zizsuk" href="#" class="showhidelink" onclick="showHide('zizsuk');return false;">\# `r auton()`</a> `r chop("The *exercises/data* folder contains a Shapefile of neighborhoods in San Francisco ([source](https://data.sfgov.org/Geographic-Locations-and-Boundaries/Analysis-Neighborhoods/p5b7-5n3h){target=\"_blank\"}). Import and plot it. How many neighborhoods are there?")`  

<div id="zizsuk" class="answer-code">
```{r import_nb, small.mar=TRUE, cache=TRUE}
library(rgdal, quietly = TRUE)
sf_nb <- rgdal::readOGR("../docs/data", "sf_neighborhoods")
plot(sf_nb, axes=T, asp=1)
nrow(sf_nb)
```

&#8658; there are `r nrow(sf_nb)` neighborhoods.

</div>

`r hashtitle("IMPORT A KML")`

<a id="qtn_wvnipb" href="#" class="showhidelink" onclick="showHide('wvnipb');return false;">\# `r auton()`</a> `r chop("The *exercises/data* folder contains a KML file with permitted cannabis dispensaries as of 2014 ([source](https://data.sfgov.org/Health-and-Social-Services/Permitted-Medical-Cannabis-Dispensaries/46p7-cb56){target=\"_blank\"}). Import and plot it over the neighborhood boundaries.")`  <a href="#" class="showhidehint" onclick="showHide('mqepye');return false;"></a>  

<div id="mqepye" class="hint">Remember to use the <tt>add=TRUE</tt> argument to overlay features on an existing plot.</div>

<div id="wvnipb" class="answer-code">
```{r plot_pot, small.mar=TRUE, cache=TRUE}
library(rgdal, quietly = TRUE)
fn <- "../docs/data/sf_permitted_cannabis_dispensaries_2014.kml"
file.exists(fn)
rgdal::ogrListLayers(fn)
sf_pot <- rgdal::readOGR(fn, "Cannabis Dispensaries 2014")
plot(sf_nb, axes=T, asp=1)
plot(sf_pot, add=T, col="red", pch=16)
```
</div>

<a id="qtn_cbaddd" href="#" class="showhidelink" onclick="showHide('cbaddd');return false;">\# `r auton()`</a> `r chop("How many permitted cannabis dispensaries are in the city?")`  

<div id="cbaddd" class="answer-code">
```{r}
nrow(sf_nb)
```
&#8658; there were `r nrow(sf_nb)` cannabis dispensaries as of 2014.
</div>


</div>

